The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-17 methods/ms.tex
distmaps

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sig$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ mJy/beam RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$35\arcsec, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

This paper serves as a companion and guide to the public data release through
NASA's Infrared Processing and Analysis Center (IPAC).  These images represent
an important finder chart for future observations with ALMA and a
long-wavelength spectral band complementary to imminent submillimeter surveys
with SCUBA-2 and Herschel-SPIRE.

@@

Millimeter-wavelength continuum surveys of the Galactic plane provide the most
efficient way to find molecular clumps that are the likely sites of formation
for massive stars and star clusters. The development of detector arrays have
made blind surveys of large areas possible.  Such surveys bypass the need for
selection based on the presence of embedded stars or star clusters, infrared
sources, masers, or radio continuum emission. In particular, they can locate
molecular clumps before stars form, providing vital information on the initial
conditions of star formation.  These surveys can also provide valuable
constraints on the physical properties of the clumps, especially masses and
mean densities.

@@

To pursue this question, we need tracers of higher density material.
Studies of nearby clouds have demonstrated that star formation is far
from uniform over molecular clouds, but is concentrated in unusually dense
regions called clumps or cores, which fill a small fraction of the area of most
clouds \citep{enoch06,evans09}[refs. E. Lada et al. Evans et al., Enoch et al.]
(Neal, John, Jonathan) Following \citet{williams00} and \citep{mckee07}, we use
the term `core' to refer to a very dense region destined to form an individual
star or small multiple star systems and the term `clump' to refer to a region
likely to form a group or cluster of stars. While clumps forming massive
clusters may be denser than cores forming low mass stars [ref Mueller? (Neal,
John, Jonathan)], in general, clumps are larger and less dense. Clumps can be
traced with line emission from molecules such as \ammonia, \form, CS, HCN, etc.
Spectral lines also provide excellent diagnostics of line-of-sight motions,
temperatures, and densities.  However, the interpretation of these gas tracers
is complicated by variations in tracer abundances caused by depletions and
complex chemical processing, uncertainties in optical depths and excitation
conditions, and the impacts of radiation fields and shocks. These complications
make the derivation of column densities, masses, and other physical properties
of star-forming clumps very difficult.  A limitation of spectral line studies
is that the tracers of denser gas tend to be weak and multi-element receiver
arrays have only a modest number of elements, making blind surveys very
time-consuming.

@@

Submillimeter observations of Galactic sources have been conducted by
a variety of different groups focusing on small regions since
bolometer technology was developed.  The Submillimeter Common-User
Bolometer Array (SCUBA) on the 15-m James Clerk Maxwell Telescope
mapped many significant regions in the galactic plane over its 8 year
lifetime \citep{holland99,difrancesco08}, but never completed a
contiguous survey of the galactic plane.  The Bolocam instrument has been
used on the 10.4-m Caltech Submiliimeter Observatory (CSO) to map nearby large
molecular clouds at 1.1 mm and identify hundreds of cores \citep{enoch06}.
Others have used MAMBO on the 30-m IRAM telescope at 1.2 mm to map
more distant regions of more massive star formation \citep{motte07}.
None of these surveys covered a significant fraction of the Galactic plane.

@@

This paper presents the results of the Bolocam Galactic Plane Survey (BGPS).
The survey covers the contiguous range $-10.5 < \lon < 86.5, -.5 < b < .5$ with
additional coverage to higher $|b|$ in the range $77 < \lon < 86.5$ and in the
$\lon=111$, W3/4/5, Gem OB1, and IC 1396 regions.  It is complementary to the
survey by \citet{schuller09} in that it targets the part of the plane visible
from the northern hemisphere and uses a somewhat longer wavelength, allowing
comparison of properties in the region of overlap with the ATLASGAL survey.
At 870 \micron, $h\nu/k = 16.5$ K, compared to 13.1 K at 1.1 mm.  Because
starless clumps may be quite cold ($\td \approx 10$ K), the uncertainty in mass
caused by uncertain temperature is slightly less at 1.1 mm.  For example, the
mass estimate if the actual $\td = 20$ K, rather than 10 K, would decrease by a
factor of 3.3 at 870 \micron\ versus a factor of 2.9 at 1.1 mm. In addition,
the Bolocam bandpass excludes the CO \jj{2}{1} line, which can be a significant
contaminant on top of the dust continuum emission.

We have used the 144-element Bolocam focal plane array, mounted at the
Cassegrain focus of the CSO, to survey over \bgpsarea\
square degrees of the northern Galactic plane at a wavelength of 1.1
mm in the dust continuum.  Figure \ref{fig:Coverage} shows the
coverage of the \bgps.  Maps of the entire BGPS are presented in
Figure \ref{fig:BGPSMontage}.  We have detected \ncores\ clumps in the
surveyed fields, providing an unbiased sample of clumps suitable for
multi-wavelength and high-resolution studies with existing telescopes
and future facilities.  At the distances that we are observing these
objects, we are likely observing aggregates of star forming cores
found in a larger molecular cloud environment, which we will refer to
as clumps \citep{williams00}.  The clump catalog is described in a
companion paper \citep{rosolowsky09}.

@@

%\Figure{mass_limits}{The limiting gas mass of cores detectable
%with at a flux limit of 15 mJy RMS. Detection thresholds at 3 (dashed)
%and 5 (solid) $\sigma$ are shown for a range of dust temperatures.
%Dust temperatures $<20$~K are most interesting for this survey, since
%these are the sources most likely to be pre- or proto-stellar
%condensations.}{fig:mass_limits}{1.0}
%
\Figure{coverage_moll} {The coverage of the BGPS.  The background
greyscale is IRAS 100 \mum.  
%The Bolocam coverage is shown with white
%solid lines in the top two panels and black in the bottom three.
%Major regions of Galactic star formation are labeled.  The coverage of
%{\em Spitzer}-GLIMPSE \citep{benjamin03} and the FCRAO $^{13}$CO GRS
%\citep{jackson06} are shown.
}{fig:Coverage}{1.0}

@@

%\begin{sidewaysfigure}[h]
\begin{figure}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$\lon=-10.5$ to $\lon=19.5$ Units are Jy/Beam.  The brightest
	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
	$\lon=13$, appear to be saturated, but this is only a display
	artifact.  The noise is more pronounced from $\lon=-7$ to $\lon=-2$
	because this region was observed less.}
    \end{center}
    \label{fig:BGPSMontage}
  \end{minipage}
\end{figure}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

@@

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

>>>>

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sigma$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ mJy/beam RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$35\arcsec, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

This paper serves as a companion and guide to the public data release through
NASA's Infrared Processing and Analysis Center (IPAC).  These images represent
an important finder chart for future observations with ALMA and a
long-wavelength spectral band complementary to ATLASGAL and imminent
submillimeter surveys with SCUBA-2 and Herschel-SPIRE.

@@

Millimeter-wavelength continuum surveys of the Galactic plane provide the most
efficient way to find molecular clumps that are the likely sites of formation
for massive stars and star clusters. The development of detector arrays have
made blind surveys of large areas possible.  Such surveys bypass the need for
selection based on the presence of embedded stars or star clusters, infrared
sources, masers, or radio continuum emission. In particular, they can locate
molecular clumps before stars form, providing vital information on the initial
conditions of star formation.  These surveys can also provide valuable
constraints on the physical properties of the clumps, especially masses and
mean densities, in conjunction with distance information.

@@

To pursue this question, we need tracers of higher density material.
Studies of nearby clouds have demonstrated that star formation is far
from uniform over molecular clouds, but is concentrated in unusually dense
regions called clumps or cores, which fill a small fraction of the area of most
clouds \citep{enoch06,evans09}[refs. E. Lada et al. Evans et al., Enoch et al.]
(Neal, John, Jonathan) Following \citet{williams00} and \citep{mckee07}, we use
the term `core' to refer to a very dense region destined to form an individual
star or small multiple star systems and the term `clump' to refer to a region
likely to form a group or cluster of stars. While clumps forming massive
clusters may be denser than cores forming low mass stars [ref Mueller? (Neal,
John, Jonathan)], in general, clumps are larger and have a lower mean density. Clumps can be
traced with line emission from molecules such as \ammonia, \form, CS, HCN, etc.
Spectral lines also provide excellent diagnostics of line-of-sight motions,
temperatures, and densities.  However, the interpretation of these gas tracers
is complicated by variations in tracer abundances caused by depletions and
complex chemical processing, uncertainties in optical depths and excitation
conditions, and the impacts of radiation fields and shocks. These complications
make the derivation of column densities, masses, and other physical properties
of star-forming clumps very difficult.  A limitation of spectral line studies
is that the tracers of denser gas tend to be weak and multi-element receiver
arrays have only a modest number of elements, making blind surveys very
time-consuming.

@@

Submillimeter observations of Galactic sources have been conducted by
a variety of different groups focusing on small regions since
bolometer technology was developed.  The Submillimeter Common-User
Bolometer Array (SCUBA) on the 15-m James Clerk Maxwell Telescope
mapped many significant regions in the galactic plane over its 8 year
lifetime \citep{holland99,difrancesco08}, but never completed a
contiguous survey of the galactic plane.  The Bolocam instrument has been
used on the 10.4-m Caltech Submiliimeter Observatory (CSO) to map nearby large
molecular clouds at 1.1 mm and identify hundreds of cores \citep{enoch06}.
MAMBO on the 30-m IRAM telescope has been used at 1.2 mm to map
more distant regions of more massive star formation \citep{motte07}.
None of these surveys covered a significant fraction of the Galactic plane.

@@

This paper presents the results of the Bolocam Galactic Plane Survey (BGPS).
The survey covers the contiguous range $-10.5 < \lon < 86.5, -.5 < b < .5$ with
additional coverage to higher $|b|$ in the range $77 < \lon < 86.5$ and in the
$\lon=111$, W3/4/5, Gem OB1, and IC 1396 regions.  It is complementary to the
survey by \citet{schuller09} in that it targets the part of the plane visible
from the northern hemisphere and uses a somewhat longer wavelength, allowing
comparison of properties in the region of overlap with the ATLASGAL survey.
At 870 \micron, $h\nu/k = 16.5$ K, compared to 13.1 K at 1.1 mm.  Because
starless clumps may be quite cold ($\td \approx 10$ K), the uncertainty in mass
caused by uncertain temperature is slightly less at 1.1 mm.  For example, 
if the actual $\td = 20$ K, rather than 10 K, the mass estimate would decrease by a
factor of 3.3 at 870 \micron\ versus a factor of 2.9 at 1.1 mm.

In addition,
the Bolocam bandpass excludes the CO \jj{2}{1} line, which can be a significant
contaminant on top of the dust continuum emission.

We have used the 144-element Bolocam focal plane array, mounted at the
Cassegrain focus of the CSO, to survey over \bgpsarea\
square degrees of the northern Galactic plane at a wavelength of 1.1
mm in the dust continuum.  Figure \ref{fig:Coverage} shows the
coverage of the \bgps.  Maps of the entire BGPS are presented in
Figure \ref{fig:BGPSMontage}.  We have detected \ncores\ clumps in the
surveyed fields, providing an unbiased sample of clumps suitable for
multi-wavelength and high-resolution studies with existing telescopes and
future facilities.  At the distances that we are likely to be observing these
objects (assuming the majority are at distances greater than 2kpc....
because?), we are likely observing aggregates of star forming cores found in a
larger molecular cloud environment, which we will refer to as clumps
\citep{williams00}.  The clump catalog is described in a companion paper
\citep{rosolowsky09}.

@@

%\Figure{mass_limits}{The limiting gas mass of cores detectable
%with at a flux limit of 15 mJy RMS. Detection thresholds at 3 (dashed)
%and 5 (solid) $\sigma$ are shown for a range of dust temperatures.
%Dust temperatures $<20$~K are most interesting for this survey, since
%these are the sources most likely to be pre- or proto-stellar
%condensations.}{fig:mass_limits}{1.0}
%
\Figure{coverage_moll} {The coverage of the BGPS.  The background
greyscale is IRAS 100 \mum.  [James, team: is there a better figure 
option?]
%The Bolocam coverage is shown with white
%solid lines in the top two panels and black in the bottom three.
%Major regions of Galactic star formation are labeled.  The coverage of
%{\em Spitzer}-GLIMPSE \citep{benjamin03} and the FCRAO $^{13}$CO GRS
%\citep{jackson06} are shown.
}{fig:Coverage}{1.0}

@@

\addtocounter{figure}{0}
%\begin{sidewaysfigure}[h]
\begin{figure}
    \label{fig:BGPSMontage}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$\lon=-10.5$ to $\lon=19.5$ Units are Jy/Beam. 
%      The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.
        The noise is more pronounced from $\lon=-7$ to $\lon=-2$
	because this region was observed less.}
    \end{center}
  \end{minipage}
\end{figure}

\addtocounter{figure}{0}
\addtocounter{subfig}{1}

@@

\addtocounter{figure}{0}
\addtocounter{subfig}{1}
